---
title: "Descriptive Statistical Analysis"
output: html_notebook
---

```{r}
library(ggplot2)
library(qqplotr)
library(tidyverse)
library(RVAideMemoire)
library(Hmisc)
```


```{r}
df <- read.csv("Gender_stats.csv", header =TRUE)
```



```{r}
head(df)
```

#DISPLAYING COLUMN NAMES
```{r}
names(df)
```

# RENAMING ALL COLUMNS NAMES
```{r}
colnames(df) <- c("Year", "Country", "Female_LifeExpectancyAtBirth", "Male_LifeExpectancyAtBirth", "Female_Population(30-34)", "Male_Population(30-34)", "Male_Population(15-19)", "Female_Population(15-19)")
```

#DISPLAYING ALL COLNAMES After the change of names

```{r}
colnames(df)
```
# to display the first six rowws
```{r}
head(df)
```

#to display the last six rows of dataframe
```{r}
tail(df)
```

```{r}
str(df)
```
# discriptive statistical analysis of each column grouped by country
```{r}
by(df, df$Country, summary)
```

#CALCULATING THE STANDARD DEVIATION

```{r}
# using lapply() function for standard deviation of all the columns at once

lapply(df[, 3:8], sd)
```

# To calculate the mode Female_lifeExpectancyAtBirth
```{r}
tab <- table(df$Female_LifeExpectancyAtBirth) # number of occurrences for each unique value
sort(tab, decreasing = TRUE) # sort highest to lowest
```
The mode of the variable Female_LifeExpectancyAtBirth is thus 81.2.

# to calculate the mode of Male_LifeExpectancyAtBirth
```{r}
tab <- table(df$Male_LifeExpectancyAtBirth) # number of occurrences for each unique value
sort(tab, decreasing = TRUE)
```
The mode of Male_LifeExpectancyAtBirth is thus 76.3

# to calculate mode of Female population (30-34)
```{r}
tab <- table(df$`Female_Population(30-34)`) # number of occurrences for each unique value
sort(tab, decreasing = TRUE) # sort from highest to lowest
```

# to calculate mode of Male population (30-14)
```{r}
tab <- table(df$`Male_Population(30-34)`) # number of occurrences for each unique value
sort(tab, decreasing = TRUE)
```

# to calculate mode of Male ppopulation (15-19)
```{r}
tab <- table(df$`Male_Population(15-19)`) # number of occurrences for each unique value
sort(tab, decreasing = TRUE) # sort highest to lowest
```

# to calculate the mode of Female population (15-19)
```{r}
tab <- table(df$`Female_Population(15-19)`) # number of occurrences for each unique value
sort(tab, decreasing = TRUE)
```

# CALCULATING THE SKEWNESS and KURTOSIS

```{r}
library(moments)
```

# CALCULATING THE SKEWNESS
```{r}
skewness(df$Female_LifeExpectancyAtBirth)
```
```{r}
skewness(df$Male_LifeExpectancyAtBirth)
```

```{r}
skewness(df$`Female_Population(30-34)`)
```

```{r}
skewness(df$`Male_Population(30-34)`)
```
```{r}
skewness(df$`Male_Population(15-19)`)
```
```{r}
skewness(df$`Female_Population(15-19)`)
```
```{r}
skewness(df$Year)
```



# CALCULATING THE KURTOSIS
```{r}
kurtosis(df$Female_LifeExpectancyAtBirth)
```

```{r}
kurtosis(df$Male_LifeExpectancyAtBirth)
```

```{r}
kurtosis(df$`Female_Population(30-34)`)
```

```{r}
kurtosis(df$`Male_Population(30-34)`)
```
```{r}
kurtosis(df$`Male_Population(15-19)`)
```

```{r}
kurtosis(df$`Female_Population(15-19)`)
```


```{r}
kurtosis(df$Year)
```


#CORELATION ANALYSIS

```{r}
# before doing correlation analysis, i have to remove the categorical varaible from the dataset.

new_df <- df %>% select(-Country, -Year)
```

```{r}
head(new_df)
```



# THE COR() FUNCTION ALLOWS US TO COMPUTE CORRELATION COEFFICIENT FOR SEVERAL PAIRS OF VARIABLES

```{r}
# correlation for all the variables
round(cor(new_df), 
      digits = 2) # rounded to 2 decimals
```



#IMPROVED CORRELATION MATRIX

```{r}
library(corrplot)
```


```{r}
corrplot(cor(new_df),
  method = "number",
  type = "upper" # show only upper side
)
```

#CORRELATION TEST FOR SEVERAL VARIABLES

*** Similar to the correlation matrix used to compute correlation for several pairs of variables, the rcorr() function (from the {Hmisc} package) allows to compute p-values of the correlation test for several pairs of variables at once. Applied to our dataset, we have:

```{r}
# correlation tests for whole dataset 
res <- rcorr(as.matrix(new_df)) # rcorr() accepts matrices only

# display p-values (rounded to 3 decimals)
round(res$P, 3)
```



