---
title: "Simple Linear Regression"
output: html_notebook
---

```{r}
library(ggplot2)
library(qqplotr)
library(tidyverse)
library(RVAideMemoire)
library(performance)
library(see)
library(car)
library(car)
```

```{r}
df <- read.csv("Gender_stats.csv", header =TRUE)
```


```{r}
head(df)
```

#DISPLAYING COLUMN NAMES
```{r}
names(df)
```

# RENAMING ALL COLUMNS NAMES
```{r}
colnames(df) <- c("Year", "Country", "Female_LifeExpectancyAtBirth", "Male_LifeExpectancyAtBirth", "Female_Population(30-34)", "Male_Population(30-34)", "Male_Population(15-19)", "Female_Population(15-19)")
```

#DISPLAYING ALL COLNAMES

```{r}
colnames(df)
```

```{r}
head(df)
```

```{r}
summary(df)
```

# seleting the numerical variables
```{r}
df_model <- df %>% select(-Year, -Country)
```

```{r}
head(df_model)
```


```{r}
# correlation between all numerical variables

round(cor(df_model), 
      digits = 3)
```
```{r}
library(corrplot)
```


```{r}
# to visualize it on a graph

corrplot(cor(df_model),
         method = "number",
         type = "upper")
```
the correlation between Female_Population(30-40) and Male_Population(15-19) is high (0.96)

#formular
Y = Female_Population(30-34)
X1 = Male_Population(15-19)


#checking for normality
```{r}
shapiro.test(df_model$`Female_Population(30-34)`)
```

```{r}
shapiro.test(df_model$`Male_Population(15-19)`)
```
```{r}
shapiro.test(df_model$Female_LifeExpectancyAtBirth)
```

# Log Transformation
```{r}
df_model$`Female_Population(30-34)` <- log10(df_model$`Female_Population(30-34)`)
```

```{r}
shapiro.test(df_model$`Female_Population(30-34)`)
```
```{r}
df_model$`Male_Population(15-19)` <- log10(df_model$`Male_Population(15-19)`)
```

```{r}
shapiro.test(df_model$`Male_Population(15-19)`)
```
```{r}
df_model$Female_LifeExpectancyAtBirth <- log10(df_model$Female_LifeExpectancyAtBirth)
```

```{r}
shapiro.test(df_model$Female_LifeExpectancyAtBirth)
```
```{r}
df_model$Male_LifeExpectancyAtBirth <- log10(df_model$Male_LifeExpectancyAtBirth)
```

```{r}
shapiro.test(df_model$Male_LifeExpectancyAtBirth)
```


so lets try and fit a SLR between Female_Population(30-34) and Female_Population(15-19)
```{r}
s_model_1 <- lm(`Female_Population(30-34)`~ `Male_Population(15-19)`, df_model)
```
```{r}
summary.lm(s_model_1)
```

```{r}
library(report)
```

```{r}
report(s_model_1)
```


```{r}
plot(`Female_Population(30-34)` ~ `Male_Population(15-19)` , df_model,
     col = "blue",
     main = "Regression: Female_Population(30-34) & Male_Population(15-19)",
     xlab = "Male_Population(15-19)",
     ylab = "Female_Population(30-34)")

```
```{r}
plot(`Female_Population(30-34)` ~ `Male_Population(15-19)` , df_model,
     col = "blue",
     main = "Regression: Female_Population(30-34) & Male_Population(15-19)",
     xlab = "Male_Population(15-19)",
     ylab = "Female_Population(30-34)")
abline(s_model_1, col = "red")
```


```{r}
plot(s_model_1, 1)
```


```{r}
plot(s_model_1, 2)
```

```{r}
plot(s_model_1, 3)
```

#Multiple Linear Regression

```{r}
corrplot(cor(df_model))
```
#Formular
Y = Female_Population(30-34)
X1 = Male_Population(15-19)
X2 = Female_LifeExpectancyAtBirth


```{r}
model_2 <- lm(`Female_Population(30-34)`~  `Male_Population(15-19)` + Female_LifeExpectancyAtBirth, df_model) 
```


```{r}
summary(model_2)
```

#checking whether the fitted model meets the MLR assumptions

i will use check_model() from {performance} package to test the assumptions at once.


```{r}
check_model(model_2)
```

```{r}
vif(model_2)
```





